CS6213 Project Report

### 2021-05-07 report/paper.tex
Apalache

    \item Hoping for a better model checking experience, we adapted our specification for use with the Apalache symbolic model checker.
    %
    This required adding type annotations to the TLA specification and slightly rewriting some operators to fit into the subset of TLA that can be automatically translated to SMT.
    %
    Apalache is alpha software, so we ran into a large number of bugs.
    %
    Sadly, for our specification, Apalache symbolic model checking was orders of magnitude slower than TLC.

@@

On several ocassions, we encountered (trivial) runtime errors after considerable model-checking time (> 15 minutes).
%
We know this is an issue in practice, and not just for us, because the specification of one-at-a-time membership changes for Raft that we adapted had two such issues which, with the default TLC options, would have taken on the order of hours to manifest at runtime.
%
This class of problem would be solved by the addition of a type system to TLA.

@@

>>>>

    \item Hoping for a better model checking experience~\cite{mcmillan_symbolic_1992}, we adapted our specification for use with the Apalache symbolic model checker.
    %
    This required adding type annotations to the TLA specification and slightly rewriting some operators to fit into the subset of TLA that can be automatically translated to SMT.
    %
    Apalache is alpha software, so we ran into a large number of bugs.
    %
    Sadly, for our specification, Apalache symbolic model checking was orders of magnitude slower than TLC.

@@

On several ocassions, we encountered (trivial) runtime errors after considerable model-checking time (> 15 minutes).
%
We know this is an issue in practice, and not just for us, because the specification of one-at-a-time membership changes for Raft that we adapted had two such issues which, with the default TLC options, would have taken on the order of hours to manifest.
%
This class of problem would be solved by the addition of a type system to TLA.

@@

The Apalache symbolic model checker~\cite{konnov_tla+_2019} does in fact add a type system to TLA.
%
Sadly, this is experimental at this stage, and we encountered several bugs related to type checking and type inference in Apalache.
%
We spent multiple days adding types to the Raft TLA specification and interpreting Apalache's obscure error messages with Igor Konnov's help, but sadly were disappointed by the symbolic model checker's performance.
%
The bad performance matches what the authors report in the paper for model checking, but we had hoped that the SMT solvers Apalache uses are more suited for goal-directed search (for test case generation) than TLC's naive breadth-first search.
%
This turned out not to be the case.

Despite our experience, we are quite hopeful about Apalache's future prospects.
%
A more intelligent translation to SMT is likely to produce good results---SMT solvers have shown themselves very capable of tackling search problems in other domains.
